---
title: "Why Quarto?"
date: "3/03/2023"
description: "Why I build my website with Quarto..."
categories: "Quarto"
format: html
editor: visual
highlight-style: github
---

## What is Quarto?

Some describe it as magic, and others have not used Quarto. I was introduced to Quarto as it was added to RStudio in an update in July 2022. Once I saw Thomas Mock's talk [Quarto for the curious](https://posit.co/talks/quarto-for-rmarkdown-users/) I knew I had to explore Quarto. I would go into detail about the technical aspects of what Quarto does, but why not show what it can do? This very website is built with Quarto, and so is [Quarto's website](https://quarto.org/).

### My website workflow

I built this site using RStuido as the environment, but if you prefer something like \`VS Code\` then you can use the Quarto [extension](https://marketplace.visualstudio.com/items?itemName=quarto.quarto). My site's code lives in it's own GitHub repository which you can view by clicking the code icon on the right side of the navigation bar...or [here](https://github.com/bradfordjohnson/quarto-portfolio). I use GitHub for version control, but also host my site with GitHub Pages. I bought my domain name and it was simple to get it up and running from there.

When I want to do some **sprucing up** to my site I create a new branch and edit the branch. Once I edit the files and code I will render my site in RStudio and then push the commits. Finally I will do a pull request and merge the branch with my main branch. GitHub Actions will then publish the site, however you do not need to do the branching. A single commit on the main branch will trigger GitHub Actions and update your site automatically. This means you can edit and publish the edits all from the **code editor**!

### Back to Quarto

Quarto documents `.qmd` are the files that you edit, and they have `YAML` code blocks at the top of the document that let you control different features such as: titles, what the output is, and many other controls. The output control is very special as you can render a file as `.html` output for webpages and many other formats! Do you need a slideshow? A `.pdf` report? A book?

Here is what the YAML looks like for this blog post!

``` yaml
---
title: "Why Quarto?"
date: "3/03/2023"
description: "Why I build my website with Quarto..."
categories: "Quarto"
format: html
editor: visual
highlight-style: github
---
```

You can create and run code chunks, this allows you to create reports, KPI dashboards, articles and so much more. I would describe Quarto as a way to share data analytics or knowledge in a meaningful and completely customized way. It is easily integrated into workflows, and you do not even have to know `R`. You can use Quarto with `Python`, `Julia`, `Observable` or `R`. Have a team of `R` users and `Python` users? They all can work together with Quarto, and even within the same document...

#### R Code Chunk

```{r}
paste0(R.Version()[c("major","minor")], collapse = ".")
```

#### Python Code Chunk

```{python}
import sys
print("Python version")
print (sys.version)
```

#### How about some code that makes a basic visual?

``` r
#| warning: false
#| message: false

library(tidyverse)

ggplot(data = iris, aes( x = Sepal.Length, y = Sepal.Width,
                         shape = Species, color = Species)) +
  geom_point()
```

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)

ggplot(data = iris, aes( x = Sepal.Length, y = Sepal.Width,
                         shape = Species, color = Species)) +
  geom_point()

```

Did you notice in the code chunk the `#|`'s? This is how you tell the chunk how to behave, I stopped warnings and messages in that chunk. You can use Quarto as **Jupyter notebooks** or **R Markdown notebooks** and render those notebooks to your desired output. Another thing is that you can use a visual editor, all within your development environment, making it that much easier to make some beautiful deliverables or notes for yourself!

## So why Quarto?

-   **Language agnostic** for data teams

-   **High quality outputs**, and **many types** of outputs

-   **Flexible workflows** and ease of use

-   It is fun, and works via **magic**

If you are still not convinced...

> I encourage you to look into Quarto as it would take a dedicated blog to really do Quarto justice, as one post can only cover so much...
>
> [Here's a Quarto blog](https://quarto.org/docs/blog/) made with Quarto, by the Quarto team!
